---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo=TRUE, warning=FALSE, results='hide', message=FALSE}
library(dplyr)
library(lattice)
activity<-read.csv('activity.csv')

```


## What is mean total number of steps taken per day?
```{r echo=TRUE}
steps.date<-activity %>% 
          filter(!is.na(steps)) %>% 
          group_by(date)  %>% 
          summarize(steps=sum(steps))

hist(steps.date$steps,main="Total number of steps taken",xlab="steps")


steps.mean<-mean(steps.date$steps)
steps.median<-median(steps.date$steps)
```  
Mean steps per day : `r steps.mean`  
Median steps per day : `r steps.median`  

## What is the average daily activity pattern?
```{r echo=TRUE}
steps.interval <- activity %>% 
               filter(!is.na(steps)) %>% 
               group_by(interval) %>% 
               summarize(steps=mean(steps))

plot(steps.interval$interval,steps.interval$steps,type="l", 
          main="Average number of steps taken per inetrval",
          ylab="steps", xlab="inteval")

steps.max<-steps.interval$interval[steps.interval$steps==max(steps.interval$steps)]
```  
The 5 minutes interval starting at `r steps.max` minutes contains the maximum evarage number of steps.


## Imputing missing value
```{r echo=TRUE}
na_count<-nrow(activity[!(complete.cases(activity)),])
```

The sourse data contains `r na_count` missing values, NA's.  
  
I replace the missing values with the culculated mean of steps per interval.
```{r echo=TRUE}
activity.na_added<-activity[!(complete.cases(activity)),c(2,3)] %>% 
                    inner_join(steps.interval,by="interval")
activity.filled<-rbind(activity[(complete.cases(activity)),],activity.na_added)
steps.filled.date<-activity.filled %>% 
                    group_by(date)  %>% 
                    summarize(steps=sum(steps))

hist(steps.filled.date$steps,main="Total number of steps taken each day, NA's substituted",xlab="steps")

steps.mean.filled<-mean(steps.filled.date$steps)
steps.median.filled<-median(steps.filled.date$steps)
```  

After subtituted the NA's we got:  
Mean steps per day: 
`r steps.mean.filled`  
Median steps per day : 
`r steps.median.filled`  

The differens between the two dataset, with and without replacing the NA's, are very small. The mean is identical and the median value has moved to be the same as mean.

## Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE, warning=FALSE, results='hide', message=FALSE}
#Change system time setting, to get english day names 
curr_locale <- Sys.getlocale("LC_TIME")
Sys.setlocale("LC_TIME", "C")

wd<-data.frame(day=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"),
               daytype=c("weekday","weekday","weekday","weekday","weekday","weekend","weekend"))

steps.typeofday<-activity.filled %>% 
               mutate(day=weekdays(as.Date(activity.filled$date))) %>% 
               inner_join(wd, by="day") %>% 
               group_by(interval,daytype) %>% 
               summarize(steps=mean(steps))

xyplot(steps ~ interval | factor(daytype), data=steps.typeofday, type="l",
     layout=c(1,2),ylab="number of steps")

#Set the system time back to original setting
Sys.setlocale("LC_TIME",curr_locale)
```  
The weekend values are more even distibuted over the day than the weekday values. The weekday values has as bigger peak at the interval between 800 and 900 minutes.